Paper 1 GLM

### 2018-10-10 Vitis_GLM.Rmd
lm attempt 1

---
title: "Paper 1: Vitis"
output: html_notebook
---

@@

```{r}
library(tidyverse)
library(tidyr)
library(skimr)
library(readxl)
read_excel("R_DiLP_all_species.xlsx")
excel_sheets("R_DiLP_all_species.xlsx")
read_excel("R_DiLP_Vitis_all_added.xlsx")
excel_sheets("R_DiLP_Vitis_all_added.xlsx")
```

# t-tests
First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth. To do this, we ran a series of t-tests.

We first tested each of the variables for *Vitis acerifolia*.

```{r}
ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
str(ac_all)
ac_all$year <- as.factor(ac_all$year)
ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
skim(ac_all)
t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.int.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$average.tooth.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.blade.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.int.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$perimeter.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$perimeter.ratio~ac_all$year, var.equal = TRUE)
t.test(ac_all$compactness~ac_all$year, var.equal = TRUE)
t.test(ac_all$shape.factor~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.blade.area~ac_all$year, var.equal = TRUE)
```
Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).

We next tested each of the variables for *Vitis aestivalis*.

```{r}
av_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "aestivalis.all.90")
av_all$year <- as.factor(av_all$year)
av_all$petiole.width.area <- as.numeric(av_all$petiole.width.area)
av_all$petiole.area <- as.numeric(av_all$petiole.area)
av_all$major.feret <- as.numeric(av_all$major.feret)
av_all$tooth.area <- as.numeric(av_all$tooth.area)
av_all$feret.diameter.ratio <- as.numeric(av_all$feret.diameter.ratio)
av_all$tooth.area.perimeter <- as.numeric(av_all$tooth.area.perimeter)
av_all$tooth.area.int.perimeter <- as.numeric(av_all$tooth.area.int.perimeter)
av_all$average.tooth.area <- as.numeric(av_all$average.tooth.area)
av_all$tooth.area.blade.area <- as.numeric(av_all$tooth.area.blade.area)
str(av_all)
skim(av_all)
t.test(av_all$total.teeth~av_all$year, var.equal = TRUE)
t.test(av_all$feret.diameter.ratio~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.int.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$average.tooth.area~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.blade.area~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.int.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$perimeter.area~av_all$year, var.equal = TRUE)
t.test(av_all$perimeter.ratio~av_all$year, var.equal = TRUE)
t.test(av_all$compactness~av_all$year, var.equal = TRUE)
t.test(av_all$shape.factor~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.blade.area~av_all$year, var.equal = TRUE)
```
Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).

We next tested each of the variables for *Vitis amurensis*.

```{r}
am_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "amurensis.all.90")
am_all$year <- as.factor(am_all$year)
am_all$petiole.width <- as.numeric(am_all$petiole.width)
am_all$petiole.width.area <- as.numeric(am_all$petiole.width.area)
am_all$petiole.area <- as.numeric(am_all$petiole.area)
am_all$blade.area <- as.numeric(am_all$blade.area)
am_all$feret.diameter <- as.numeric(am_all$feret.diameter)
am_all$major.feret <- as.numeric(am_all$major.feret)
am_all$leaf.area <- as.numeric(am_all$leaf.area)
am_all$perimeter <- as.numeric(am_all$perimeter)
am_all$internal.perimeter <- as.numeric(am_all$internal.perimeter)
am_all$feret.diameter.ratio <- as.numeric(am_all$feret.diameter.ratio)
am_all$tooth.area.perimeter <- as.numeric(am_all$tooth.area.perimeter)
am_all$tooth.area.int.perimeter <- as.numeric(am_all$tooth.area.int.perimeter)
am_all$tooth.area.blade.area <- as.numeric(am_all$tooth.area.blade.area)
am_all$teeth.perimeter <- as.numeric(am_all$teeth.perimeter)
am_all$teeth.int.perimeter <- as.numeric(am_all$teeth.int.perimeter)
am_all$perimeter.area <- as.numeric(am_all$perimeter.area)
am_all$perimeter.ratio <- as.numeric(am_all$perimeter.ratio)
am_all$compactness <- as.numeric(am_all$compactness)
am_all$shape.factor <- as.numeric(am_all$shape.factor)
am_all$teeth.blade.area <- as.numeric(am_all$teeth.blade.area)
str(am_all)
skim(am_all)
t.test(am_all$total.teeth~am_all$year, var.equal = TRUE)
t.test(am_all$feret.diameter.ratio~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.int.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$average.tooth.area~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.blade.area~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.int.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$perimeter.area~am_all$year, var.equal = TRUE)
t.test(am_all$perimeter.ratio~am_all$year, var.equal = TRUE)
t.test(am_all$compactness~am_all$year, var.equal = TRUE)
t.test(am_all$shape.factor~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.blade.area~am_all$year, var.equal = TRUE)
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).

We finally tested each of the variables for *Vitis riparia*.

```{r}
r_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "riparia.all.90")
r_all$year <- as.factor(r_all$year)
r_all$petiole.width <- as.numeric(r_all$petiole.width)
r_all$petiole.width.area <- as.numeric(r_all$petiole.width.area)
r_all$petiole.area <- as.numeric(r_all$petiole.area)
r_all$internal.perimeter <- as.numeric(r_all$internal.perimeter)
r_all$tooth.area <- as.numeric(r_all$tooth.area)
r_all$tooth.area.perimeter <- as.numeric(r_all$tooth.area.perimeter)
r_all$tooth.area.int.perimeter <- as.numeric(r_all$tooth.area.int.perimeter)
r_all$average.tooth.area <- as.numeric(r_all$average.tooth.area)
r_all$tooth.area.blade.area <- as.numeric(r_all$tooth.area.blade.area)
r_all$teeth.int.perimeter <- as.numeric(r_all$teeth.int.perimeter)
r_all$perimeter.ratio <- as.numeric(r_all$perimeter.ratio)
r_all$teeth.blade.area <- as.numeric(r_all$teeth.blade.area)
str(r_all)
skim(r_all)
t.test(r_all$total.teeth~r_all$year, var.equal = TRUE)
t.test(r_all$feret.diameter.ratio~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.int.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$average.tooth.area~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.blade.area~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.int.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$perimeter.area~r_all$year, var.equal = TRUE)
t.test(r_all$perimeter.ratio~r_all$year, var.equal = TRUE)
t.test(r_all$compactness~r_all$year, var.equal = TRUE)
t.test(r_all$shape.factor~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.blade.area~r_all$year, var.equal = TRUE)
```

Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).

# Principal Components Analysis (PCA)

```{r}
library(pcaMethods)
excel_sheets("R_DiLP_Vitis_all_added.xlsx")
```

First, we applied a PCA to *Vitis acerifolia*.
```{r}
ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
ac_ln_fixed <- ac_ln_mat[,-(1:3)]
ac_ln_fixed[ac_ln_fixed == 0] <- NA
#pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_ac_ll)
loadings(bpca_ac_ll)
#summary(pca_ac_all)
#loadings(pca_ac_all)
print(bpca_ac_ll)
#print(pca_ac_all)
#slplot(pca_ac_all)
#loadings_pca_ac <- loadings(pca_ac_all)
#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
#str(loadings_pca_ac_df)
#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
  #geom_col(x = PC1, y = PC2) +
  #ggtitle("PCA plot of Vitis acerifolia")
#attach row names
```

We then applied a PCA to *Vitis aestivalis*.

```{r}
av_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "aestivalis.all.ln")
av_ln_mat <- data.matrix(av_ln, rownames.force = NA)
av_ln_fixed <- av_ln_mat[,-(1:3)]
av_ln_fixed[av_ln_fixed == 0] <- NA
#pca_av_all <- pca(av_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_av_ll <- pca(av_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_av_ll)
loadings(bpca_av_ll)
#summary(pca_av_all)
#loadings(pca_av_all)
print(bpca_av_ll)
#print(pca_av_all)
#slplot(pca_av_all)
```

We then applied a PCA to *Vitis amurensis*.

```{r}
am_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "amurensis.all.ln")
am_ln_mat <- data.matrix(am_ln, rownames.force = NA)
am_ln_fixed <- am_ln_mat[,-(1:3)]
am_ln_fixed[am_ln_fixed == 0] <- NA
#pca_am_all <- pca(am_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_am_ll <- pca(am_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_am_ll)
loadings(bpca_am_ll)
#summary(pca_am_all)
#loadings(pca_am_all)
print(bpca_am_ll)
#print(pca_am_all)
#slplot(pca_am_all)
```

Finally, we applied a PCA to *Vitis riparia*.

```{r}
r_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "riparia.all.ln")
r_ln_mat <- data.matrix(r_ln, rownames.force = NA)
r_ln_fixed <- r_ln_mat[,-(1:3)]
r_ln_fixed[r_ln_fixed == 0] <- NA
#pca_r_all <- pca(r_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_r_ll <- pca(r_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_r_ll)
loadings(bpca_r_ll)
#summary(pca_r_all)
#loadings(pca_r_all)
print(bpca_r_ll)
#print(pca_r_all)
#slplot(pca_r_all)
```

# Generalized Linear Modeling

@@

We ran a GLM for leaf number and total teeth.

```{r}
skim(ac_all)
ac_tt_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
summary(ac_tt_glm)
plot(ac_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
ac_fdr_glm <- glm(leaf~feret.diameter.ratio, data = ac_all, family = gaussian())
summary(ac_fdr_glm)
plot(ac_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
ac_tap_glm <- glm(leaf~tooth.area.perimeter, data = ac_all, family = gaussian())
summary(ac_tap_glm)
plot(ac_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
ac_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = ac_all, family = gaussian())
summary(ac_taip_glm)
plot(ac_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
ac_ata_glm <- glm(leaf~average.tooth.area, data = ac_all, family = gaussian())
summary(ac_ata_glm)
plot(ac_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
ac_taba_glm <- glm(leaf~tooth.area.blade.area, data = ac_all, family = gaussian())
summary(ac_taba_glm)
plot(ac_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
ac_tp_glm <- glm(leaf~teeth.perimeter, data = ac_all, family = gaussian())
summary(ac_tp_glm)
plot(ac_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
ac_tip_glm <- glm(leaf~teeth.int.perimeter, data = ac_all, family = gaussian())
summary(ac_tip_glm)
plot(ac_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
ac_pa_glm <- glm(leaf~perimeter.area, data = ac_all, family = gaussian())
summary(ac_pa_glm)
plot(ac_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
ac_pr_glm <- glm(leaf~perimeter.ratio, data = ac_all, family = gaussian())
summary(ac_pr_glm)
plot(ac_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
ac_cp_glm <- glm(leaf~compactness, data = ac_all, family = gaussian())
summary(ac_cp_glm)
plot(ac_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
ac_sf_glm <- glm(leaf~shape.factor, data = ac_all, family = gaussian())
summary(ac_sf_glm)
plot(ac_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
ac_tba_glm <- glm(leaf~teeth.blade.area, data = ac_all, family = gaussian())
summary(ac_tba_glm)
plot(ac_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
ac_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = ac_all, family = gaussian())
summary(ac_clim_glm)
plot(ac_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
ac_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = ac_all, family = gaussian())
summary(ac_all_glm)
plot(ac_all_glm)
```

We tested *Vitis aestivalis*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(av_all)
av_tt_glm <- glm(leaf~total.teeth, data = av_all, family = gaussian())
summary(av_tt_glm)
plot(av_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
av_fdr_glm <- glm(leaf~feret.diameter.ratio, data = av_all, family = gaussian())
summary(av_fdr_glm)
plot(av_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
av_tap_glm <- glm(leaf~tooth.area.perimeter, data = av_all, family = gaussian())
summary(av_tap_glm)
plot(av_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
av_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = av_all, family = gaussian())
summary(av_taip_glm)
plot(av_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
av_ata_glm <- glm(leaf~average.tooth.area, data = av_all, family = gaussian())
summary(av_ata_glm)
plot(av_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
av_taba_glm <- glm(leaf~tooth.area.blade.area, data = av_all, family = gaussian())
summary(av_taba_glm)
plot(av_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
av_tp_glm <- glm(leaf~teeth.perimeter, data = av_all, family = gaussian())
summary(av_tp_glm)
plot(av_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
av_tip_glm <- glm(leaf~teeth.int.perimeter, data = av_all, family = gaussian())
summary(av_tip_glm)
plot(av_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
av_pa_glm <- glm(leaf~perimeter.area, data = av_all, family = gaussian())
summary(av_pa_glm)
plot(av_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
av_pr_glm <- glm(leaf~perimeter.ratio, data = av_all, family = gaussian())
summary(av_pr_glm)
plot(av_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
av_cp_glm <- glm(leaf~compactness, data = av_all, family = gaussian())
summary(av_cp_glm)
plot(av_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
av_sf_glm <- glm(leaf~shape.factor, data = av_all, family = gaussian())
summary(av_sf_glm)
plot(av_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
av_tba_glm <- glm(leaf~teeth.blade.area, data = av_all, family = gaussian())
summary(av_tba_glm)
plot(av_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
av_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = av_all, family = gaussian())
summary(av_clim_glm)
plot(av_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
av_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = av_all, family = gaussian())
summary(av_all_glm)
plot(av_all_glm)
```

we tested *Vitis amurensis*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(am_all)
am_tt_glm <- glm(leaf~total.teeth, data = am_all, family = gaussian())
summary(am_tt_glm)
plot(am_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
am_fdr_glm <- glm(leaf~feret.diameter.ratio, data = am_all, family = gaussian())
summary(am_fdr_glm)
plot(am_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
am_tap_glm <- glm(leaf~tooth.area.perimeter, data = am_all, family = gaussian())
summary(am_tap_glm)
plot(am_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
am_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = am_all, family = gaussian())
summary(am_taip_glm)
plot(am_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
am_ata_glm <- glm(leaf~average.tooth.area, data = am_all, family = gaussian())
summary(am_ata_glm)
plot(am_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
am_taba_glm <- glm(leaf~tooth.area.blade.area, data = am_all, family = gaussian())
summary(am_taba_glm)
plot(am_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
am_tp_glm <- glm(leaf~teeth.perimeter, data = am_all, family = gaussian())
summary(am_tp_glm)
plot(am_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
am_tip_glm <- glm(leaf~teeth.int.perimeter, data = am_all, family = gaussian())
summary(am_tip_glm)
plot(am_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
am_pa_glm <- glm(leaf~perimeter.area, data = am_all, family = gaussian())
summary(am_pa_glm)
plot(am_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
am_pr_glm <- glm(leaf~perimeter.ratio, data = am_all, family = gaussian())
summary(am_pr_glm)
plot(am_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
am_cp_glm <- glm(leaf~compactness, data = am_all, family = gaussian())
summary(am_cp_glm)
plot(am_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
am_sf_glm <- glm(leaf~shape.factor, data = am_all, family = gaussian())
summary(am_sf_glm)
plot(am_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
am_tba_glm <- glm(leaf~teeth.blade.area, data = am_all, family = gaussian())
summary(am_tba_glm)
plot(am_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
am_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = am_all, family = gaussian())
summary(am_clim_glm)
plot(am_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
am_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = am_all, family = gaussian())
summary(am_all_glm)
plot(am_all_glm)
```

We tested *Vitis riparia*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(r_all)
r_tt_glm <- glm(leaf~total.teeth, data = r_all, family = gaussian())
summary(r_tt_glm)
plot(r_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
r_fdr_glm <- glm(leaf~feret.diameter.ratio, data = r_all, family = gaussian())
summary(r_fdr_glm)
plot(r_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
r_tap_glm <- glm(leaf~tooth.area.perimeter, data = r_all, family = gaussian())
summary(r_tap_glm)
plot(r_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
r_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = r_all, family = gaussian())
summary(r_taip_glm)
plot(r_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
r_ata_glm <- glm(leaf~average.tooth.area, data = r_all, family = gaussian())
summary(r_ata_glm)
plot(r_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area.

```{r}
r_taba_glm <- glm(leaf~tooth.area.blade.area, data = r_all, family = gaussian())
summary(r_taba_glm)
plot(r_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
r_tp_glm <- glm(leaf~teeth.perimeter, data = r_all, family = gaussian())
summary(r_tp_glm)
plot(r_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
r_tip_glm <- glm(leaf~teeth.int.perimeter, data = r_all, family = gaussian())
summary(r_tip_glm)
plot(r_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
r_pa_glm <- glm(leaf~perimeter.area, data = r_all, family = gaussian())
summary(r_pa_glm)
plot(r_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
r_pr_glm <- glm(leaf~perimeter.ratio, data = r_all, family = gaussian())
summary(r_pr_glm)
plot(r_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
r_cp_glm <- glm(leaf~compactness, data = r_all, family = gaussian())
summary(r_cp_glm)
plot(r_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
r_sf_glm <- glm(leaf~shape.factor, data = r_all, family = gaussian())
summary(r_sf_glm)
plot(r_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
r_tba_glm <- glm(leaf~teeth.blade.area, data = r_all, family = gaussian())
summary(r_tba_glm)
plot(r_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
r_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = r_all, family = gaussian())
summary(r_clim_glm)
plot(r_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
r_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = r_all, family = gaussian())
summary(r_all_glm)
plot(r_all_glm)
beep(sound = 3, expr = NULL)
```

>>>>

---
title: "Paper 1 GLM"
output: html_notebook
---

@@

```{r setup}
library(readxl)
library(dplyr)
library(skimr)
library(ggplot2)
library(segmented)
```

# Linear Modeling

@@

We ran a LM for leaf number and total teeth.
```{r}
acerifolia <-filter(R_DiLP_all_species, species == "acerifolia")
skim(acerifolia, total.teeth)
ac_tt_plot <- ggplot(acerifolia, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic()
ac_tt_plot <- ac_tt_plot +labs(x = "Leaf number",
                               y = "Total teeth")
ac_tt_plot

```

```{r}
ac_tt_lm <- lm(total.teeth ~ leaf, data = acerifolia)
summary(ac_tt_lm)
```

```{r}
ac_tt_seg <- segmented(ac_tt_lm,
                       seg.Z = ~ leaf,
                       psi = NA)
summary(ac_tt_seg)
```

First, we tested *Vitis aestivalis*.

We ran a GLM for leaf number and total teeth.
```{r}
aestivalis <-filter(R_DiLP_all_species, species == "aestivalis")
```

First, we tested *Vitis amurensis*.

We ran a GLM for leaf number and total teeth.
```{r}
amurensis <-filter(R_DiLP_all_species, species == "amurensis")
```

First, we tested *Vitis riparia*.

We ran a GLM for leaf number and total teeth.
```{r}
riparia <-filter(R_DiLP_all_species, species == "riparia")
```
